---
title: "Case Study 1"
subtitle: "Group 97 - R Programming"
author:
  - Beccard\, Leon - 12133103
  - Farmohammadi\, Pouya
  - Monteiro Oliveira\, Guilherme - 12243299
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: True
    toc_float:
      collapsed: true
      smooth_scroll: true
    toc_depth: 1
    theme: readable
    # number_sections: True
editor_options: 
  markdown: 
    wrap: 80
---

--------------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Ratio of Fibonacci numbers

## a.

Write two different R functions which return the sequence ri = Fi+1 /Fi for i =
1, . . . , n where Fi is the ith Fibonacci number, once using for and once using
while.

For understanding the sequence, it is as follows\
Fibonacci sequence: 0,1,1,2,3,5,8,13,21,..., with 0 as F0

-   r1 = F2/F1 = 1/1 = 1\
-   r2 = F3/F2 = 2/1 = 2\
-   r3 = F4/F3 = 3/2 = 1.5

**For loop**

```{r fib_ratio_for}
fib_ratio_for <- function(n) {
  # Check for bad inputs
  if (n <= 0) {
    stop("Input 'n' cannot be 0 or negative.")
  } else if (n == 1) {
    # Catch case of n = 1 due to a leading error if not implemented
    return(1)
  } else {
    # numeric vector created with length n + 1 
    # to store the Fibonacci sequence up to the (n + 1)th element.
    fibonacci <- numeric(n + 1) 
    # Assign values (1,1) as first two elements
    fibonacci[1:2] <- c(1, 1)
    # for loop that iterates over the sequence 3:(n + 1)
    # starting from 3 because first two elements are already declared
    for (i in 3:(n + 1)) {
      # Assigning Fibonacci rule to get the next number out of the last two
      fibonacci[i] <- fibonacci[i - 1] + fibonacci[i - 2]
    }
    # Finally calculate the ratio of each fibonacci number
    ratios <- fibonacci[2:(n + 1)] / fibonacci[1:n]
    return(ratios)
  }
}
```

Execute for loop:

```{r fib_ratio_for(5)}
fib_ratio_for(5)
```

**While loop**

```{r fib_ratio_while}
fib_ratio_while <- function(n) {
  if (n <= 0) {
    stop("Input 'n' cannot be 0 or negative.")
  } else if (n == 1) {
    return(1)
  } else {
    fibonacci <- numeric(n + 1)
    fibonacci[1:2] <- c(1, 1)
    i <- 3
    while (i <= n + 1) {
      fibonacci[i] <- fibonacci[i - 1] + fibonacci[i - 2]
      i <- i + 1
    }
    ratios <- fibonacci[2:(n + 1)] / fibonacci[1:n]
    return(ratios)
  }
}
```

Execute while loop:

```{r fib_ratio_while(7)}
fib_ratio_while(7)
```

## b.

Benchmark the two functions for n = 200 and n = 2000 (you can use package
microbenchmark or package bench for this purpose). Which function is faster?

```{r benchmark}
# install.packages('microbenchmark')
library(microbenchmark)

microbenchmark(
  fib_ratio_for(200),
  fib_ratio_while(200),
  fib_ratio_for(2000),
  fib_ratio_while(2000)
)
```

**Explanation for statistics given from microbenchmark**\
min: Minimum time it took to execute\
lq: Lower quartile (25th percentile) time it took to execute\
mean: Mean time it took to execute\
median: Median time it took to execute\
uq: Upper quartile (75th percentile) time it took to execute\
max: Maximum time it took to execute\
neval: Number of times each expression was evaluated

**Answer:** As given from the stats, the function using the for loop is faster.

## c.

Plot the sequence for n = 100. For which value n it starts to converge?\
What is the number that it converges to?

```{r plot}
# Calculate the sequence for n = 100
sequence <- fib_ratio_for(100)

# Plot the sequence
plot(sequence, type = "l", xlab = "Index (i)", ylab = "Ratio (r_i)",
     main = "Sequence of Fibonacci Ratios for n = 100")

```

```{r tail_sequence}
# Get the last 6 values to know what number it converges to
tail(sequence)
```

**Answer:** It converges towards the 1.618034. The so-called golden ratio.

# 2. Gamma function

## a.

## b.

## c.

## d.

# 3. The golden ration

## a.

## b.

## c.

# 4. Game of craps

A variation of game of craps is played as follows. The player rolls a die and if
they roll a 6, they immediately wins. If a 1 is rolled, they immediately loses.
If a 2, 3, 4, or 5 is rolled, then this number becomes their “point number.’ ’
The player then has up to three more rolls of the die to roll their point number
again. If they do, they win otherwise they lose. Write a program in R to
simulate a game of craps. Explain the steps of your program. If the code is not
explained, no points will be earned.

```{r game_of_crabs}
# Define the function for a game of crabs
game_of_craps <- function(){
    # Function to get a random number between 1 and 6 i.e. rolling the dice once
    dice_number <- floor(runif(1, min=1, max=7))
    # check for win or lose condition in first attempt
    if (dice_number == 1){
      return("Sorry, you lost!")
    }
    else if (dice_number == 6){
      return("Yeah, you won in the first attempt!")
    }
    else {
      # for loop to have three more attempts to roll the same 'point number'
      for (i in 1:3){
        # Function to get a random number between 1 and 6 i.e. rolling the dice once
        new_dice_number <- floor(runif(1, min=1, max=7))
        # Check for same 'point number'
        if (dice_number == new_dice_number){
          return("Yeah, you won rolling the same 'point number' again!")
        }
    }
  } 
  return("Sorry, you lost after three more attempts!")
}

# Execute function i.e. one simulation of the game
game_of_craps()

```

--------------------------------------------------------------------------------

# 5. Readable and efficient code

Read over the code below.

```{r redable, results='hide'}

set.seed(1)
x <- rnorm(1000)
y <- 2 + x + rnorm(1000)
df <- data.frame(x, y)

cat("Step", 1, "\n")
fit1 <- lm(y ~ x, data = df[-(1:250),])
p1 <- predict(fit1, newdata = df[(1:250),])
r <- sqrt(mean((p1 - df[(1:250),"y"])^2))

cat("Step", 2, "\n")
fit2 <- lm(y ~ x, data = df[-(251:500),])
p2 <- predict(fit2, newdata = df[(251:500),])
r <- c(r, sqrt(mean((p2 - df[(251:500),"y"])^2)))

cat("Step", 3, "\n")
fit3 <- lm(y ~ x, data = df[-(501:750),])
p3 <- predict(fit3, newdata = df[(501:750),])
r <- c(r, sqrt(mean((p3 - df[(501:750),"y"])^2)))

cat("Step", 4, "\n")
fit4 <- lm(y ~ x, data = df[-(751:1000),])
p4 <- predict(fit4, newdata = df[(751:1000),])
r <- c(r, sqrt(mean((p4 - df[(751:1000),"y"])^2)))
r

```


## Solution {.tabset}

### a. Explain the code

The code initially generates an random variable x defined by a normal distribution 
with mean  and standard deviation of 1. Then, it generates another *noise* variable with the same characteristics as x and y is a combination of x, the noise and a constant of 2. 

Each step consist of r selecting different sections of the data frame df with a proportion of 
75% to 25% for training and testing respectively and calculate the root mean square error for each step.
The error for each steps are `r r` and the biggest gap between the errors is  `r max(r) - min(r)`, what is doesn't looks like outstanding.


### b. Readability

From the code we can notice a few things:

* The code is repeating itself a few times and run it in a loop or call functions to avoid code repetition would make it more readable. 
* The code could write more comments in the code to give us an abstraction of what he is planning to do and inline comment to give us specific details, like the definition of the variable x.
* Use meaningful names can make it easier to understand the code, like *pred1*, instead of *p1*.


### c. Wrap Function

Change the code according to a. and wrap it in a function. This function should have at most 10 lines
(without adding commands to more lines such as x <- 1; y <- 2. Such commands will count as 2 lines!).
Check that the function called on the same input outputs the same as the provided code. 

```{r efficient_5, results='hold'}

# Define the function

linear_models <- function(df, formula, steps = 4) {
  # Create an empty vector to store the errors
  r <- numeric()
  step = nrow(df)/steps
  # Loop through the steps
  for (i in 1:steps) {
    cat("Step", i, "\n")
    section <- ((i-1)*step+1):(i*step)
    fit <- lm(formula, data = df[-section,]) # Fit the model
    p <- predict(fit, newdata = df[section,]) # Predict the model
    r <- c(r, sqrt(mean((p - df[section,"y"])^2)))
  }
  return(r)
}

set.seed(1)
x <- rnorm(1000)
y <- 2 + x + rnorm(1000)
df <- data.frame(x, y)

# Execute the function

linear_models(df, y ~ x)

```


--------------------------------------------------------------------------------

# 6. Measuring and improving performance

Have a look at the code of the function below. It is a function for performing a
Kruskal Wallis test, a robust non-parametric method for testing whether samples
come from the same distribution. (Note: we assume no missing values are present
in x).

```{r Kruskal_Wallis_test}

kwtest <- function (x, g, ...)
{
  if (is.list(x)) {
    if (length(x) < 2L)
      stop("'x' must be a list with at least 2 elements")
    if (!missing(g))
      warning("'x' is a list, so ignoring argument 'g'")
    if (!all(sapply(x, is.numeric)))
      warning("some elements of 'x' are not numeric and will be coerced to numeric")
    k <- length(x)
    l <- lengths(x)
    if (any(l == 0L))
      stop("all groups must contain data")
    g <- factor(rep.int(seq_len(k), l))
    x <- unlist(x)
}
else {
  if (length(x) != length(g))
    stop("'x' and 'g' must have the same length")
  g <- factor(g)
  k <- nlevels(g)
  if (k < 2L)
    stop("all observations are in the same group")
}
n <- length(x)
if (n < 2L)
  stop("not enough observations")
r <- rank(x)
TIES <- table(x)

STATISTIC <- sum(tapply(r, g, sum)^2/tapply(r, g, length))
STATISTIC <- ((12 * STATISTIC/(n * (n + 1)) - 3 * (n + 1))/(1 -
  sum(TIES^3 - TIES)/(n^3 - n)))
PARAMETER <- k - 1L

PVAL <- pchisq(STATISTIC, PARAMETER, lower.tail = FALSE)
names(STATISTIC) <- "Kruskal-Wallis chi-squared"
names(PARAMETER) <- "df"
RVAL <- list(statistic = STATISTIC, parameter = PARAMETER,
  p.value = PVAL, method = "Kruskal-Wallis rank sum test")
return(RVAL)
}

```

## Solution {.tabset}

### **a. Pseudo code**

This is a test

```{r pseudo_code, eval=FALSE}

kwtest_pseudo <- function (x, g, ...)
{
  if x is a list {
    if x has less than 2 elements
      Do not continue with the code and inform the user
    if g is provided
      Give a warning to the user that g will be ignored
    if all the elemensts of x are not numeric
      Warns the users that some elements are not numeric and will be converted 
      to numeric
    k <- length(x)
    l <- lengths(x)
    if any element of x is 0
      Stop and inform that all the groups should contain data
    g <- factor(rep.int(seq_len(k), l))
    x <- unlist(x)
  }
  if x is not a list {
    if the length of x and g are different
      Stops and inform the user x and g should be the same length
    g <- factor(g)
    k <- nlevels(g)
    if there are less than 2 groups in g
      Stop and inform the user
  }
  n <- length(x)
  if x length is less than 2
    Stop and inform about the lack of observations
  r <- rank(x)
  TIES <- table(x)
  STATISTIC <- Calculate the Kruskal-Wallis statistic
  PARAMETER <- degree of freedom
  PVAL <- pchisq(STATISTIC, PARAMETER, lower.tail = FALSE)
  names(STATISTIC) <- "Kruskal-Wallis chi-squared"
  names(PARAMETER) <- "df"
  RVAL <- list(statistic = STATISTIC, parameter = PARAMETER,
  p.value = PVAL, method = "Kruskal-Wallis rank sum test")
  return(RVAL)
  
}

```

### **b. Example call**

```{r example_call, results="hold"}

# Example with list

x_list <- list(
  group1 = c(1, 3, 4, 2),
  group2 = c(2, 6, 1, 7)
)

# Example with vector

x_vector <- c(1, 3, 4, 2, 2, 6, 1, 7)
g_vector <- c(1, 1, 1, 1, 2, 2, 2, 2) # group labels

result_list <- kwtest(x_list)
result_vector <- kwtest(x_vector, g_vector)

print("Result using x as list:")
print(result_list)

print("Result using x as vector:")
print(result_vector)


```


### **c. kwtest faster**

```{r kwtest_faster, results='hold'}

kwtest_fast <- function(x, g, ...) {
  if (!is.numeric(x) | !is.numeric(g)) {
    stop("x and g must be numeric vectors")
  }
  
  if (length(x) != length(g)) {
    stop("x and g must have the same length")
  }
  
  #if (nlevels(g) < 2)
   # print(paste("G", g))
    #print(paste("nlevels", nlevels(g)))
    #stop("nlevels(g) < 2 all observations are in the same group")
  
  if (n < 2L)
    stop("not enough observations")
  
  N <- length(x)
  number_groups <- nlevels(factor(g))
  obs_group <- table(g)
  r <- rank(x)
  r_group_avg <- tapply(r, g, sum)/obs_group
  r_avg <- (N + 1)/2
  
  H_num <-(N - 1) * sum(obs_group * (r_group_avg - r_avg)**2)
  H_dem <- sum((r - r_avg)**2)
  
  H <- H_num/H_dem
  
}

```

### d. Multiple experiments

```{r}
set.seed(1234)
m <- 1000 # number of repetitions
n <- 50
# number of individuals
X <- matrix(rt(m * n, df = 10), nrow = m)
grp <- rep(1:3, c(20, 20, 10))


kwtest_matrix <- function(X, grp) {
  res <- apply(X, 1, function(x) kwtest(x, grp))
}

# Returns all the statistics from res

res <- kwtest_matrix(X, grp)

# Extract the p-values

p_values <- sapply(res, function(x) x$p.value)


```


### e.
```{r}
kwtest_matrix2 <- function(X, g) {
  if (!is.matrix(X)) {
    stop("X must be a matrix.")
  }
  if (!is.numeric(g)) {
    stop("g must be a numeric vector")
  }
  if (ncol(X) != length(g)) {
    stop("Number of cols X of matrix have to be as long as length of vector g")
  }
  m <- nrow(X)
  test_statistics <- numeric(m)
  
  for (i in 1:m) {
    # Take data from each row
    experiment_data <- X[i,]
    # Calculate Chi-Square value from Kruskal wallis test and store it in vector
    test_statistics[i] = kwtest_fast(experiment_data, g)
  }
  
  return(test_statistics)
  
}


```
### f.
```{r}
library(microbenchmark)

# For a better overview set random data from above here again
set.seed(1234)
m <- 1000 # number of repetitions
n <- 50
X <- matrix(rt(m * n, df = 10), nrow = m)
grp <- rep(1:3, c(20, 20, 10))

# Benchmark the two functions
benchmarks <- microbenchmark(
  apply_approach = kwtest_matrix(X, grp),
  loop_approach = kwtest_matrix2(X, grp),
  times = 10
)

# Print the benchmarks
print(benchmarks)

```


### g.
