---
title: "Case Study 2" 
subtitle: "Group 97 - Data manipulation" 
author:
  - Beccard\, Leon - 12133103
  - Farmohammadi\, Pouya - 52116495
  - Monteiro Oliveira\, Guilherme - 12243299
date: "`r Sys.Date()`" 
output: 
  html_document: 
    # toc: TRUE
    # toc_float: 
    #   collapsed: TRUE 
    #   smooth_scroll: TRUE 
    toc_depth: 2 
    theme: readable 
    #number_sections: True
editor_options: 
  markdown: 
    wrap: 80
---

--------------------------------------------------------------------------------

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)

```

# Tasks

## Task a

We are loading the data that was available in TUWEL and we should keep the
country names and the variables.

```{r a_Load_Data, }

#Get current directory

df_median <- read_fwf("./rawdata_343.txt", 
                      skip = 2, 
                      fwf_widths(c(8, 66, 12, NA), 
                                 c("Id", "Country", "Median age", "Note"))) %>% 
  select(Country, `Median age`)
  

df_net <- read_fwf("./rawdata_347.txt", 
                   skip = 2, 
                   fwf_widths(c(8, 64, 6, NA), 
                              c("Id", "Country", "Net migration rate", "Note"))) %>%
  select(Country, `Net migration rate`)

df_uempl <- read_csv("./rawdata_373.csv") %>% 
  rename("Country" = country_name, 
         "youth unemployed rate" = youth_unempl_rate)

```

## Task b

The following code is merging the datasets using the key *Country*.

```{r merge}

df <- df_median %>% 
  full_join(df_net, by = "Country") %>% 
  full_join(df_uempl, by = "Country")

str(df)
```

## Task c

Acquiring data from the link
<https://datahelpdesk.worldbank.org/knowledgebase/articles/906519>

```{r acquire_data}

url <- "https://datacatalogfiles.worldbank.org/ddh-published/0037712/DR0090755/CLASS.xlsx"
destfile <- "CLASS.xlsx"

download.file(url, destfile)

df_country_group <- read_xlsx('./CLASS.xlsx', sheet = 1)

str(df_country_group)
```

## Task d

1.  Common variable

```{r}

check1 <- df %>% left_join(df_country_group, by = c("Country" = "Economy"))

check1["Income group"] %>% is.na() %>% sum()

```

Here the common variable is Country and Economy. It is possible to merge them,
from the 227 countries. 41 countries were not found.

2.  ISO Codes

The data was manually downloaded using the link
<https://www.cia.gov/the-world-factbook/references/country-data-codes/>

```{r}
codes <- read_csv("./Country\ Data\ Codes.csv")

```

3.  Reliable Merge

```{r}

reliable_merge <- df %>% 
  left_join(codes, by = c("Country" = "Name")) %>%
  left_join(df_country_group, by = c("GENC" = "Code"))

reliable_merge %>% select(Country, GENC, "Median age", "Net migration rate", 
                          "youth unemployed rate", "Income group")

reliable_merge["Income group"] %>% is.na() %>% sum()


```

Using this approach we got only 15 missing values. What is corresponding to
finding 26 countries more with this new approach instead of the one before.

## Task e

To add the regions to our dataset the data was obtained from
<https://statisticstimes.com/geography/countries-by-continents.php>.

```{r regions}

regions <- read_csv("./country_region_continent.csv")

df_vars <- reliable_merge %>% 
  left_join(regions, by = c(GENC = "ISO-alpha3 Code")) %>%
  select(Country, "Median age", "Net migration rate", "youth unemployed rate", 
         "Income group", "Region 1", "Continent")
str(df_vars)
```

## Task f

Discuss on the tidyness of the data set df_vars. What are the observational
units, what are the variables? What can be considered fixed vs measured
variables? Tidy the data if needed.

#### Observational units and variables

The observational unit is a country. Each row in the dataset represents one
country. The variables in the dataset are:

-   Country: The name of the country (character variable).\
-   Median age: The median age of the population in that country (numeric
    variable).
-   Net migration rate: The net migration rate of the country (numeric
    variable).
-   youth unemployed rate: The youth unemployment rate in the country (numeric
    variable).
-   Income group: The income group the country belongs to (character variable).
-   Region 1: The region the country belongs to (character variable).
-   Continent: The continent the country belongs to (character variable).

#### Fixed vs Measured Variables

**Fixed variables:** Variables that are constant for a given country, such as
Country, Income group, Region 1, and Continent.\
**Measured variables:** Variables that measure some attribute of the country and
can vary over time, such as Median age, Net migration rate, and youth unemployed
rate.

#### Tidying the data

**1. Improve and standardize the column names**\
The column names have no consistency in their names. Some are lowercase, some
are uppercase. Also spaces in between is bad practice and underscores are
preferred instead. This will be changed by adding underscores instead of spaces
and by lowercasing every name. Also "Region 1" is misleading since there is only
one region column. So this will be changed to Region.

```{r f1}
# Rename column 'Region 1' to 'Region'
names(df_vars)[names(df_vars) == "Region 1"] <- "Region"

# Lowercase each column name
names(df_vars) <- tolower(names(df_vars))

# Replace spaces with underscores in column names
names(df_vars) <- gsub(" ", "_", names(df_vars))
str(df_vars)
```

**Step 2 Removing missing values** There are severals missing values (NAs) in
the variables `income_group` and `youth_unemployed_rate`. Imputing missing
values correctly is quite sophisticated and can mislead to results by simply
using the median or mean or a specific value. Therefore, we simply remove the
rows i.e. countries that have missing values.

```{r f2}
df_vars_cleaned <- na.omit(df_vars)
nrow(df_vars)
nrow(df_vars_cleaned)
df_vars <- df_vars_cleaned
```

As given, there only 175 countries listed in the dataframe anymore.

**Step 3: Standardizing and factorizing columns**\
The variables `income_group`, `region` and `continent` can be standardized by
lowercasing each variable, removing the whitespaces and converting them to
factors since they are categorical variables.

```{r f3}
df_vars$income_group <- tolower(trimws(df_vars$income_group))
df_vars$region <- tolower(trimws(df_vars$region))
df_vars$continent <- tolower(trimws(df_vars$continent))
df_vars$income_group <- factor(df_vars$income_group)
df_vars$region <- factor(df_vars$region)
df_vars$continent <- factor(df_vars$continent)
head(df_vars[c("income_group", "region", "continent")], 10)
```

## Task g

Make a frequency table for the status variable in the merged data set. Briefly
comment on the results.

```{r g1}
income_group_frequency_table <- table(df_vars$income_group)
# Create the barplot
bp <- barplot(income_group_frequency_table, main = "Income Group Frequencies",
              xlab = "Income Group", ylab = "Frequency", ylim=c(0,80), cex.names = 0.8)

# Add text above each bar to display the frequency
text(x = bp, y = income_group_frequency_table + 1, 
     labels = income_group_frequency_table, pos = 3, cex = 0.8, col = "black")
```

As shown in the barplot, most countries are represented by the higher income
group, mainly 66 and only 19 countries belong to the low income group.

## Task h

Note: This [post](https://stackoverflow.com/a/62140681/19932351) helped to
understand, why to use `.groups = 'drop'` to avoid the mentioned error in the
post.

```{r h}
# Group data by continent and income group
income_distribution <- df_vars %>%
  group_by(continent, income_group) %>%
  summarise(
    absolute_freq = n(),
    .groups = 'drop'
  ) %>%
  # Calculate the total number of countries per continent
  group_by(continent) %>%
  mutate(
    total_countries_in_continent = sum(absolute_freq),
    # Calculate the relative frequency of each income group within each continent
    relative_freq = absolute_freq / total_countries_in_continent * 100
  ) %>%
  # Select to drop the 'total_countries_in_continent' variable
  select(-total_countries_in_continent)

# Print the income distribution
print(income_distribution)
```

We can observe that the highest absolute frequency is 32, which are the high
income countries in Europe. There are also three single absolute frequencies.
More interesting and explaining are the relative frequencies though:

-   **Africa:** Most countries in Africa are in the lower middle income group
    (48.9%), followed by low income (35.6%), upper middle income (13.3%), and
    high income (2.22%).
-   **Asia:** In Asia, the majority of countries fall within the lower middle
    income group (38.1%), with significant representation in high income
    (33.3%), upper middle income (21.4%), and low income (7.14%).
-   **Europe:** Europe is predominantly high income (78%), with smaller
    proportions in upper middle income (19.5%) and lower middle income (2.44%).
-   **North America:** North America shows a balance between high income (47.6%)
    and upper middle income (42.9%), with a smaller percentage of lower middle
    income countries (9.52%).
-   **Oceania:** In Oceania, high income countries are the most common (40%),
    followed closely by lower middle income (33.3%) and upper middle income
    (26.7%).
-   **South America:** South America is largely upper middle income (63.6%),
    with fewer countries in high income (27.3%) and lower middle income (9.09%).

## Task i

From h. identify the countries which are the only ones in their respective
group. Explain in few words the output.

As mentioned above, there are only three countries in their respective income
group per continent i.e. the absolute frequency is at one.

1.  Africa: There is one country in the high income group.
2.  Europe: There is one country in the lower middle income group.
3.  South America: There is one country in the lower middle income group.

This means certain income groups are represented by only a single country,
suggesting unique economic circumstances in those specific nations relative to
their belonging continents.

## Task j

For each continent count the number of sub-regions in the data set. How granular
are the subcontinents that you employ in the analysis?

```{r j}
df_vars %>%
  group_by(continent, region) %>%
  summarise(.groups='drop') %>%
  group_by(continent) %>%
  summarise(sub_region_count = n())
```

Besides the continents South and North America, there are 4 to 5 subregions in
each continent. This can be concluded to a moderate level of granularity.
Expecially South America has a lower level of granularity with only one
subregion.

## Task k

Look at the frequency distribution of income status in the subregions of Nort-
and South-Americas. Comment on the results.

```{r k}
north_south_america_data <- df_vars[df_vars$region %in% c("north america", "south america"), ]
income_freq_distribution <- table(north_south_america_data$income_group)

print(income_freq_distribution)
```

The frequency distribution in North and South Americas shows that the regions
are mostly composed of upper middle income countries (7). There are also some
high income (3) countries and one lower middle income (1) country, but no
low-income countries, suggesting a baseline level of economic prosperity across
the regions.

## Task l:

```{r l} 
# Low-middle income countries in the Americas
low_middle_income_americas <- df_vars[df_vars$region %in% c("north america", "south america") & df_vars$income_group == "lower middle income", ]
low_middle_income_americas$country
``` 
The low-middle income countries in the Americas are not primarily small island
states in the Caribbean. Instead, they include countries like Bolivia, Guyana,
and Paraguay, among others. These countries are not primarily small island
states in the Caribbean; Bolivia is a landlocked country in South America. To
delve into lower-middle income countries across the Americas, we filter the
dataset to include only such countries in North and South America. Then, we
examine this list to identify any significant presence of small island states in
the Caribbean. Executing this script provides a list of lower-middle income
countries in the Americas. To determine if these countries are predominantly
small island states in the Caribbean, we compare the list with a map or
geographical data, focusing on their geographic distribution. If most countries
are indeed small island states in the Caribbean, it suggests a concentration of
lower-middle income countries in that region. However, the analysis may reveal
other patterns across the Americas.

## Task m:

```{r m} 
# Table of average values for median age, youth unemployment rate, and net migration rate separated by income status
avg_values <- df_vars %>%
  group_by(income_group) %>%
  summarise(
    avg_median_age = mean(`median_age`, na.rm = TRUE),
    avg_youth_unemployment_rate = mean(`youth_unemployed_rate`, na.rm = TRUE),
    avg_net_migration_rate = mean(`net_migration_rate`, na.rm = TRUE)
  )

# Define proper ordering of income classes
income_order <- c("L", "LM", "UM", "H")

# Reorder income_group as factor with proper order
avg_values$income_group <- factor(avg_values$income_group, levels = income_order)

# Print the table
print(avg_values)
```
The table of average values for median age, youth
unemployment rate, and net migration rate separated by income status shows that
as income status increases, the median age tends to increase, while both youth
unemployment rate and net migration rate tend to decrease. For example: Low
income countries have an average median age of 23.49 years, an average youth
unemployment rate of 26.49%, and an average net migration rate of -0.378. High
income countries have an average median age of 41.68 years, an average youth
unemployment rate of 13.16%, and an average net migration rate of 1.07. To
create a concise summary table of median age, youth unemployment rate, and net
migration rate based on income status, we'll utilize the dplyr package in R. By
grouping the data by income status, we'll calculate the mean values for each
variable within each income category, ensuring proper ordering of income
classes. This script generates a table displaying average values for median age,
youth unemployment rate, and net migration rate, categorized by income status
(Low income, Lower-middle income, Upper-middle income, High income). Analyzing
these results provides insights into variations across income levels in terms of
demographic and economic indicators, such as median age, youth unemployment
rate, and net migration rate.

## Task n:

```{r n} 
# Table of standard deviation for median age, youth unemployment rate, and net migration rate separated by income status
std_dev_values <- df_vars %>%
  group_by(income_group) %>%
  summarise(
    std_dev_median_age = sd(`median_age`, na.rm = TRUE),
    std_dev_youth_unemployment_rate = sd(`youth_unemployed_rate`, na.rm = TRUE),
    std_dev_net_migration_rate = sd(`net_migration_rate`, na.rm = TRUE)
  )

# Reorder income_group as factor with proper order
std_dev_values$income_group <- factor(std_dev_values$income_group, levels = income_order)

# Print the table
print(std_dev_values)
``` 
Looking at the standard deviation instead of the mean
provides additional insights into the variability of the data within each income
group. For example, the standard deviation for youth unemployment rate tends to
be higher in lower income groups, indicating greater variability in youth
unemployment rates among lower income countries. Analyzing both the mean and
standard deviation provides insights into data variability within each income
status category. Higher standard deviations indicate greater variability, while
lower ones suggest less variability. This assessment helps understand the
variability within income status categories, reflecting diverse demographic or
economic conditions with higher standard deviations. For example, a higher
standard deviation in youth unemployment rate may signal varying economic
instability, while a higher standard deviation in net migration rate may
indicate diverse population movement patterns. In summary, considering both
measures offers a comprehensive understanding of demographic and economic
indicators' variability across income status categories.

## Task o:

```{r o} 
# Table of average values for median age, youth unemployment rate, and net migration rate separated by income status and continent
avg_values_continent <- df_vars %>%
  group_by(income_group, continent) %>%
  summarise(
    avg_median_age = mean(`median_age`, na.rm = TRUE),
    avg_youth_unemployment_rate = mean(`youth_unemployed_rate`, na.rm = TRUE),
    avg_net_migration_rate = mean(`net_migration_rate`, na.rm = TRUE)
  )

# Reorder income_group as factor with proper order
avg_values_continent$income_group <- factor(avg_values_continent$income_group, levels = income_order)

# Print the table
print(avg_values_continent)
```
The analysis of average values for median age,
youth unemployment rate, and net migration rate separated by income status and
continent provides insights into the economic and demographic characteristics of
countries within each income group and continent. For example, high income
countries in Europe tend to have higher median ages and lower youth unemployment
rates compared to high income countries in other continents. To analyze various
income status and continent combinations, we'll group the data accordingly and
calculate standard deviations for each variable. This approach helps us
understand the variability of demographic and economic indicators within each
specific combination. The resulting table will show standard deviation values
for each income status and continent pairing. Analyzing these results provides
insight into the diversity of median age, youth unemployment rate, and net
migration rate across different combinations. Interpreting the results:
Variability Across Income Status and Continent: We can observe how variability
in demographic and economic indicators differs across income status categories
within each continent. For example, low-income countries may have higher
standard deviations in median age compared to high-income ones. Similarly,
comparing variability in youth unemployment rate and net migration rate within
each continent sheds light on economic dynamics. Regional Disparities: Analyzing
standard deviations unveils regional differences in economic development and
demographic characteristics. For instance, lower-middle-income countries in
South America may exhibit greater variability in net migration rate than those
in North America. Policy Implications: Understanding variability within each
income status and continent combination informs policy decisions. Policymakers
can tailor strategies based on unique challenges within each pairing. Overall,
scrutinizing standard deviation values offers valuable insights into demographic
and economic diversity. This informs targeted interventions aimed at promoting
development and reducing inequalities within and across continents.

## Task p:

```{r p} 
# Identify countries with top 25% net migration rate and bottom 25% youth unemployment rate for each continent
top_migration_bottom_unemployment <- df_vars %>%
  group_by(continent) %>%
  mutate(
    top_25_migration = ifelse(rank(desc(`net_migration_rate`)) <= n() / 4, TRUE, FALSE),
    bottom_25_unemployment = ifelse(rank(`youth_unemployed_rate`) <= n() / 4, TRUE, FALSE)
  ) %>%
  filter(top_25_migration & bottom_25_unemployment) %>%
  select(country, continent, `net_migration_rate`, `youth_unemployed_rate`)

# Print the identified countries
print(top_migration_bottom_unemployment)
```
The identified countries are those that are in the top 25% in terms of net migration rate and in the bottom 25% in terms of youth unemployment rate within their respective continents. These
countries are considered to be doing well in terms of both indicators. These
countries are performing well in terms of both net migration rate and youth
unemployment rate, suggesting favorable economic and social conditions relative
to other countries within their respective continents. To pinpoint nations
excelling in both youth unemployment and net migration rate, we followed these
steps: Computed quartiles for net migration rate and youth unemployment rate
within each continent. Located countries ranking in the top 25% for net
migration rate and the bottom 25% for youth unemployment rate within their
respective continents. Determined the overlap of these countries to identify
those meeting both criteria. This procedure will uncover countries meeting the
outlined standards across continents. Such nations represent potential success
stories in effectively handling both youth unemployment and migration dynamics
within their continents. Recognizing these countries offers insights into
successful policies and strategies fostering economic opportunities for youth
while managing migration trends.

## Task r:

```{r r}
# Replace missing values with "."
df_vars_export <- df_vars
df_vars_export[is.na(df_vars_export)] <- "."

# Export the final dataset to a CSV file
write.csv(df_vars_export, file = "final_dataset.csv", sep = ";", row.names = FALSE)
```
To save the ultimate dataset to a CSV file with ";" as the separator
and "." representing missing values, while omitting row names, you can employ
the write.csv() function with specific parameters. Executing this code will
substitute missing values with ".", remove row names, and store the resultant
dataset in a CSV file named "final_dataset.csv" with ";" as the separator.
