---
title: "Case Study 2" subtitle: "Group 97 - Data manipulation" author:
  - Beccard\, Leon - 12133103
  - Farmohammadi\, Pouya - 52116495
  - Monteiro Oliveira\, Guilherme - 12243299
date: "`r Sys.Date()`" output: html_document: toc: True toc_float: collapsed:
true smooth_scroll: true toc_depth: 1 theme: readable # number_sections: True
editor_options: markdown: wrap: 80
---
--------------------------------------------------------------------------------
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
--------------------------------------------------------------------------------
# 6. Measuring and improving performance
Have a look at the code of the function below. It is a function for performing a
Kruskal Wallis test, a robust non-parametric method for testing whether samples
come from the same distribution. (Note: we assume no missing values are present
in x).
```{r Kruskal_Wallis_test}
kwtest <- function (x, g, ...)
{
  if (is.list(x)) {
    if (length(x) < 2L)
      stop("'x' must be a list with at least 2 elements")
    if (!missing(g))
      warning("'x' is a list, so ignoring argument 'g'")
    if (!all(sapply(x, is.numeric)))
      warning("some elements of 'x' are not numeric and will be coerced to numeric")
    k <- length(x)
    l <- lengths(x)
    if (any(l == 0L))
      stop("all groups must contain data")
    g <- factor(rep.int(seq_len(k), l))
    x <- unlist(x)
}
else {
  if (length(x) != length(g))
    stop("'x' and 'g' must have the same length")
  g <- factor(g)
  k <- nlevels(g)
  if (k < 2L)
    stop("all observations are in the same group")
}
n <- length(x)
if (n < 2L)
  stop("not enough observations")
r <- rank(x)
TIES <- table(x)
STATISTIC <- sum(tapply(r, g, sum)^2/tapply(r, g, length))
STATISTIC <- ((12 * STATISTIC/(n * (n + 1)) - 3 * (n + 1))/(1 -
  sum(TIES^3 - TIES)/(n^3 - n)))
PARAMETER <- k - 1L
PVAL <- pchisq(STATISTIC, PARAMETER, lower.tail = FALSE)
names(STATISTIC) <- "Kruskal-Wallis chi-squared"
names(PARAMETER) <- "df"
RVAL <- list(statistic = STATISTIC, parameter = PARAMETER,
  p.value = PVAL, method = "Kruskal-Wallis rank sum test")
return(RVAL)
}
```
## Solution {.tabset}
### **a. Pseudo code**
This is a test
```{r pseudo_code, eval=FALSE}
kwtest_pseudo <- function (x, g, ...)
{
  if x is a list {
    if x has less than 2 elements
      Do not continue with the code and inform the user
    if g is provided
      Give a warning to the user that g will be ignored
    if all the elemensts of x are not numeric
      Warns the users that some elements are not numeric and will be converted
      to numeric
    k <- length(x)
    l <- lengths(x)
    if any element of x is 0
      Stop and inform that all the groups should contain data
    g <- factor(rep.int(seq_len(k), l))
    x <- unlist(x)
  }
  if x is not a list {
    if the length of x and g are different
      Stops and inform the user x and g should be the same length
    g <- factor(g)
    k <- nlevels(g)
    if there are less than 2 groups in g
      Stop and inform the user
  }
  n <- length(x)
  if x length is less than 2
    Stop and inform about the lack of observations
  r <- rank(x)
  TIES <- table(x)
  STATISTIC <- Calculate the Kruskal-Wallis statistic
  PARAMETER <- degree of freedom
  PVAL <- pchisq(STATISTIC, PARAMETER, lower.tail = FALSE)
  names(STATISTIC) <- "Kruskal-Wallis chi-squared"
  names(PARAMETER) <- "df"
  RVAL <- list(statistic = STATISTIC, parameter = PARAMETER,
  p.value = PVAL, method = "Kruskal-Wallis rank sum test")
  return(RVAL)
}
```
### **b. Example call**
```{r example_call, results="hold"}
# Example with list
x_list <- list(
  group1 = c(1, 3, 4, 2),
  group2 = c(2, 6, 1, 7)
)
# Example with vector
x_vector <- c(1, 3, 4, 2, 2, 6, 1, 7)
g_vector <- c(1, 1, 1, 1, 2, 2, 2, 2) # group labels
result_list <- kwtest(x_list)
result_vector <- kwtest(x_vector, g_vector)
print("Result using x as list:")
print(result_list)
print("Result using x as vector:")
print(result_vector)
```
### **c. kwtest faster**
```{r kwtest_faster, results='hold'}
kwtest_fast <- function(x, g, ...) {
  if (!is.numeric(x) | !is.numeric(g)) {
    stop("x and g must be numeric vectors")
  }
  if (length(x) != length(g)) {
    stop("x and g must have the same length")
  }
  number_groups <- nlevels(factor(g))
  N <- length(x)
  if (number_groups < 2L){
    stop("nlevels(g) < 2 shows all observations are in the same group")
  }
  if (N < 2L)
    stop("length(x) has not enough observations")
  number_groups <- nlevels(factor(g))
  obs_group <- length(unique(g))
  r <- rank(x)
  r_group_avg <- tapply(r, g, sum)/obs_group
  r_avg <- (N + 1)/2
  H_num <-(N - 1) * sum(obs_group * (r_group_avg - r_avg)**2)
  H_dem <- sum((r - r_avg)**2)
  H <- H_num/H_dem
  pval <- pchisq(H, obs_group - 1, lower.tail = FALSE)
  return(pval)
}
kwtest_fast(x_vector, g_vector)
```
### **d. Multiple experiments using apply function**
```{r 6d_multiple_experiments}
set.seed(1234)
m <- 1000 # number of repetitions
n <- 50
# number of individuals
X <- matrix(rt(m * n, df = 10), nrow = m)
grp <- rep(1:3, c(20, 20, 10))
kwtest_mult <- function(X, grp) {
  res <- apply(X, 1, function(x) kwtest_fast(x, grp))
}
# Returns all the statistics from res
res <- kwtest_mult(X, grp)
# Extract the p-values
p_values <- sapply(res, function(x) x)
```
### **e. Multiple experiments using for loop**
```{r 6e_multiple_experiments_for_loop}
kwtest_matrix2 <- function(X, g) {
  if (!is.matrix(X)) {
    stop("X must be a matrix.")
  }
  if (!is.numeric(g)) {
    stop("g must be a numeric vector")
  }
  if (ncol(X) != length(g)) {
    stop("Number of cols X of matrix have to be as long as length of vector g")
  }
  m <- nrow(X)
  test_statistics <- numeric(m)
  for (i in 1:m) {
    # Take data from each row
    experiment_data <- X[i,]
    # Calculate Chi-Square value from Kruskal wallis test and store it in vector
    test_statistics[i] = kwtest_fast(experiment_data, g)
  }
  return(test_statistics)
}
```
### **f. Benchmarking**
```{r 6f_benchmarking, echo = T, results = 'hide'}
library(microbenchmark)
# For a better overview set random data from above here again
set.seed(1234)
m <- 1000 # number of repetitions
n <- 50
X <- matrix(rt(m * n, df = 10), nrow = m)
grp <- rep(1:3, c(20, 20, 10))
# Benchmark the two functions
benchmarks <- microbenchmark(
  apply_approach = kwtest_mult(X, grp),
  loop_approach = kwtest_matrix2(X, grp),
  times = 10
)
# Print the benchmarks
print(benchmarks)
```
As shown from the statistics, the loop approach is faster than the apply
approach.
### **g. Vectorizing**
```{r 6g_vectorizing, echo = T, results = 'hide'}
library(microbenchmark)
# For a better overview set random data from above here again
set.seed(1234)
m <- 1000 # number of repetitions
n <- 50
X <- matrix(rt(m * n, df = 10), nrow = m)
grp <- rep(1:3, c(20, 20, 10))
kwtest_vectorize <- function(X, grp) {
  #?Vectorize
  #print(grp)
  # vectorized_kwtest <- Vectorize(kwtest_fast, vectorize.args = "X")
  #print(vectorized_kwtest)
  # res <- vectorized_kwtest(X, grp)
  # return(res)
  res <- mapply(kwtest, split(X, seq(nrow(X))), MoreArgs = list(grp))
  return(res)
}
# Benchmark the two functions
benchmarks <- microbenchmark(
  apply_approach = kwtest_mult(X, grp),
  loop_approach = kwtest_matrix2(X, grp),
  vectorize_approach = kwtest_vectorize(X, grp),
  times = 10
)
# Print the benchmarks
print(benchmarks)
```
Since the Vectorize() uses mapply() under the hood, we can use this function
right away to compare the results with the other two approaches. As shown, the
loop approach is still the best.